The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2016-02-15 script/en/chapters/crypto2020.tex
Script[de+en]: Small additions like research about Benford's law and primes.
fixed overfull \hbox and undefined references
Corrected typos and overfull hbox; congruent index (no redundancy in wording; adjectives in lowercase in German; all normal words in lowercase in English)

\newpage
\hypertarget{Chapter_Crypto2020}{}
\chapter{Crypto 2020 --- Perspectives for long-term cryptographic
         security}\index{Security!long-term}
\label{Chapter_Crypto2020}
\begin{sloppypar}
(Johannes Buchmann, Erik Dahmen, Alexander May and Ulrich Vollmer, TU Darmstadt, May~2007)\\
\end{sloppypar}

@@

In 2005, that is 27 years after the invention of
RSA, Bahr, Boehm, Franke, and Kleinjung from
Bochum University managed to factor a 200 digit
RSA challenge number
(\url{www.mat.uniroma2.it/~eal/rsa640.txt}).  A
key with size originally thought to be secure for
a very long time was broken with a computation
that took them just five months.  This illustrates
the tremendous progress factoring technology has
made within the last 30 years.  This progress is
based on break-through mathematical ideas --- e.g.\
the number field sieve proposed by John
Pollard --- as well as significant developments in
computer hardware and software implementation
technology.\footnote{%
Please compare chapter \ref{SecurityRSA}
\hyperlink{SecurityRSA}{Considerations regarding the
security of the RSA algorithm}, and especially chapters
\ref{NoteFactorization} and \ref{FactorizationResearch}.
}

In 2000, Lenstra and Verheul\index{Lenstra/Verheul}
developed an extrapolation formula that is supposed
to help us forecast the security\index{Security!forecast}
one can achieve with RSA and other important
cryptographic schemes in the long
term (\url{www.keylength.com}).  The formula
suggests the use of 850 digit RSA moduli if one
wishes to protect data for the next 30 years.
This corresponds to a 3072 bit RSA key.

Yet, even a well thought out extrapolation formula
is no security guarantee!  At any time, a
brilliant mathematical idea can allow us to factor
large numbers easily, and destroy the security of
RSA.  In 1996, Peter Shor showed that a quantum
computer --- a new type of computer that leverages
the laws of quantum mechanics to speed up certain
types of computations --- can in principle be used
for the fast factorization of large numbers \cite{shor:1997}.
Despite intensive research in the area, it is
still too early to judge whether we are ever going
to be able to build quantum computers\index{Quantum computer} of
sufficient capacity to apply Shor's algorithm to
numbers of relevant size.\footnote{%
Required qbits for attacks on RSA, DSA and ECDSA using key with a bit length n: \\
\vskip +1 pt
\begin{tabular}{|c|l|}
\hline
   RSA		&  2n + 3 \\
   DSA		&  2n + 3 \\
   ECDSA $2^n$	&  \~{}2n + 8 log n \\
   ECDSA p	&  \~{}4n \\
\hline
\end{tabular}
\vskip +6 pt
Please compare chapter 5.3 in
``SicAri -- Eine Sicherheitsplattform und deren Werkzeuge
f\"ur die ubiquit\"are Internetnutzung, KB2.1 -- Abschlussbericht,
\"Ubersicht \"uber Angriffe auf relevante kryptographische Verfahren'',
version 1.0, Mai 17, 2005,
Prof. Dr. Johannes Buchmann et al., TUD-KryptC and cv cryptovision GmbH
(\href{http://www.cdc.informatik.tu-darmstadt.de/~schepers/kb\_21\_angriffe.pdf}
 {\tt http://www.cdc.informatik.tu-darmstadt.de/\~{}schepers/kb\_21\_angriffe.pdf}) and the dissertation of Axel Schmidt at the same faculty.
}
Recent announcements of
significant progress in this area made by the
start-up company D-Wave (\url{www.dwavesys.com})
have been greeted with a lot of scepticism, even
ridicule.

@@

For many years, the cryptography group at the
Technische Universit\"at Darmstadt and its spin-off,
the company FlexSecure (\url{www.flexsecure.de}),
have worked to provide these tools.  The
trust center software FlexiTrust which is employed
by the German National Root Certification
Authority and the German Country Signing Authority
offers an infrastructure within which
cryptographic schemes can be easily exchanged.
The open source library FlexiProvider\index{FlexiProvider} implements a
multitude of cryptographic schemes.  Lately, we
have intensified our research into ``Post Quantum
Cryptography''\index{Cryptography!Post Quantum} seeking
cryptographic schemes which
remain secure even in the event that powerful
quantum computers are built.

The security of public key cryptography
traditionally rests on the difficulty of the
solution of certain mathematical problems.  Today,
the following alternatives to the factorization and discrete
logarithm problems are discussed in depth:  the
decoding problem, the shortest and closest vector
problem in lattices, and the problem of solving
large systems of multivariate quadratic
equations.  It is conjectured that quantum
computers\index{Quantum computer} offer little advantage if we try to
solve these problems efficiently.

@@

Let us look at these alternatives a little more
closely.  The first encryption scheme based on the
decoding problem was proposed by
McEliece~\cite{mceliece:1978}\index{Encryption!McEliece}.  The background:
Error-correcting codes are used to transmit or
store electronic data in such a way that they
remain undistorted even if a small number of bits
are changed in transit or on the storage media.
This property is used in, e.g., compact discs
(CDs).  The data on a CD can be reconstructed even
if the disc has been slightly scratched.

In a code-based encryption\index{Encryption!code-based} scheme a message is
encrypted by adding a fixed number of errors to
(i.e. flipping a fixed numbers of bits of) the
encoded message.  Decoding requires knowledge of a
suitable decoding procedure which eliminates these
errors efficiently.  This method is the secret
key.  Code-based encryption is in general very
efficient.  At the moment, research focus on the
question which codes lead to secure encryption
schemes with keys which are as small as possible.

Encryption on the basis of lattice problems\index{Encryption!lattice problems}
is very similar to that on the basis of
error-correcting codes.  Lattices are regular
structures of points in space.  For instance, the
points where the lines on squared paper cross form
a two-dimensional lattice.  For cryptographic
usage, the dimension of the lattices is chosen to
be much larger.  Encryption works as follows: The
plain-text is used to construct a lattice point
which is then slightly distorted in such a way
that it is no longer a lattice point, but close to
one.  Whoever knows a secret about the lattice is
able to find this lattice point in the vicinity of
the given point in space.  The lattice point in
turn yields the plain text.  A particularly
efficient lattice based encryption scheme is NTRU
Encrypt (\url{www.ntru.com})\index{Encryption!lattice problems!NTR}.
However, because
NTRU was introduced fairly recently (in 1998), and
its specification underwent several changes due to
a variety of attacks, more cryptanalytic scrutiny
is required to achieve confidence in its security.

@@

In 1979, Ralph Merkle proposed a remarkable
framework for new signature schemes in his PhD
thesis~\cite{merkle-thesis:1979}.
Contrary to all other signature schemes\index{Signature!Merkle}, its
security does not rest on the difficulty of a
number-theoretic, algebraic or geometric problem.
The only thing it requires is something which
other signature schemes need anyway: a
cryptographically secure hash function and a
secure pseudo-random number generator.  Each new
hash function leads to a new signature algorithm.
In consequence, the Merkle scheme has the
potential to solve the problem of long-term
availability of digital signature schemes.

@@

%\subsection{Quantum cryptography -- a loophole?}
\section{Quantum cryptography -- a way out of the impasse?}
\label{sec:quantum cryptography}\index{Quantum cryptography}

@@

In order to guarantee IT security for the future, too,
we need to prepare a portfolio of secure cryptographic
schemes.  This portfolio needs to contain schemes
which are suitable for the world of ubiquitous
computing with many less powerful computers.  It
also needs to contain schemes which remain secure
in the event that powerful quantum computers\index{Quantum computer} are
built.  Several promising candidates have been
discussed in this article.  They need to be
studied carefully and prepared for use in everyday
scenarios.  The question how to ensure long-term
confidentiality remains an important open research
problem upon which cryptographic research should
focus.

>>>>

\newpage
\hypertarget{Chapter_Crypto2020}{}
\chapter{Crypto 2020 --- Perspectives for Long-Term Cryptographic
         Security}\index{security!long-term}
\label{Chapter_Crypto2020}
\begin{sloppypar}
(Johannes Buchmann, Erik Dahmen, Alexander May and Ulrich Vollmer, TU Darmstadt, May~2007)\\
\end{sloppypar}

@@

In 2005, that is 27 years after the invention of
RSA, Bahr, Boehm, Franke, and Kleinjung from
Bochum University managed to factor a 200 digit
RSA challenge number
(\url{www.mat.uniroma2.it/~eal/rsa640.txt}).  A
key with size originally thought to be secure for
a very long time was broken with a computation
that took them just five months.  This illustrates
the tremendous progress factoring technology has
made within the last 30 years.  This progress is
based on break-through mathematical ideas --- e.g.\
the number field sieve proposed by John
Pollard --- as well as significant developments in
computer hardware and software implementation
technology.\footnote{%
Please compare chapter \ref{SecurityRSA}
\hyperlink{SecurityRSA}{Considerations regarding the
security of the RSA algorithm}, and especially chapters
\ref{nt:NoteFactorization} and \ref{FactorizationResearch}.
}

In 2000, Lenstra and Verheul\index{Lenstra/Verheul}
developed an extrapolation formula that is supposed
to help us forecast the security\index{security!forecast}
one can achieve with RSA and other important
cryptographic schemes in the long
term (\url{www.keylength.com}).  The formula
suggests the use of 850 digit RSA moduli if one
wishes to protect data for the next 30 years.
This corresponds to a 3072 bit RSA key.

Yet, even a well thought out extrapolation formula
is no security guarantee!  At any time, a
brilliant mathematical idea can allow us to factor
large numbers easily, and destroy the security of
RSA.  In 1996, Peter Shor showed that a quantum
computer --- a new type of computer that leverages
the laws of quantum mechanics to speed up certain
types of computations --- can in principle be used
for the fast factorization of large numbers \cite{shor:1997}.
Despite intensive research in the area, it is
still too early to judge whether we are ever going
to be able to build quantum computers\index{quantum computer} of
sufficient capacity to apply Shor's algorithm to
numbers of relevant size.\footnote{%
Required qbits for attacks on RSA, DSA and ECDSA using key with a bit length n: \\
\vskip +1 pt
\begin{tabular}{|c|l|}
\hline
   RSA		&  2n + 3 \\
   DSA		&  2n + 3 \\
   ECDSA $2^n$	&  \~{}2n + 8 log n \\
   ECDSA p	&  \~{}4n \\
\hline
\end{tabular}
\vskip +6 pt
Please compare chapter 5.3 in
``SicAri -- Eine Sicherheitsplattform und deren Werkzeuge
f\"ur die ubiquit\"are Internetnutzung, KB2.1 -- Abschlussbericht,
\"Ubersicht \"uber Angriffe auf relevante kryptographische Verfahren'',
version 1.0, Mai 17, 2005,
Prof. Dr. Johannes Buchmann et al., TUD-KryptC and cv cryptovision GmbH
(\href{http://www.cdc.informatik.tu-darmstadt.de/~schepers/kb\_21\_angriffe.pdf}
 {\tt http://www.cdc.informatik.tu-darmstadt.de/\~{}schepers/kb\_21\_angriffe.pdf}) and the dissertation of Axel Schmidt at the same faculty.
}
Recent announcements of
significant progress in this area made by the
start-up company D-Wave (\url{www.dwavesys.com})
have been greeted with a lot of scepticism, even
ridicule.

@@

For many years, the cryptography group at the
Technische Universit\"at Darmstadt and its spin-off,
the company FlexSecure (\url{www.flexsecure.de}),
have worked to provide these tools.  The
trust center software FlexiTrust which is employed
by the German National Root Certification
Authority and the German Country Signing Authority
offers an infrastructure within which
cryptographic schemes can be easily exchanged.
The open source library FlexiProvider\index{FlexiProvider} implements a
multitude of cryptographic schemes.  Lately, we
have intensified our research into ``Post Quantum
Cryptography''\index{cryptography!post quantum}\index{post-quantum computing}\index{PQC} seeking
cryptographic schemes which
remain secure even in the event that powerful
quantum computers are built.

The security of public key cryptography
traditionally rests on the difficulty of the
solution of certain mathematical problems.  Today,
the following alternatives to the factorization and discrete
logarithm problems are discussed in depth:  the
decoding problem, the shortest and closest vector
problem in lattices, and the problem of solving
large systems of multivariate quadratic
equations.  It is conjectured that quantum
computers\index{quantum computer} offer little advantage if we try to
solve these problems efficiently.

@@

Let us look at these alternatives a little more
closely.  The first encryption scheme based on the
decoding problem was proposed by
McEliece~\cite{mceliece:1978}\index{encryption!McEliece}.  The background:
Error-correcting codes are used to transmit or
store electronic data in such a way that they
remain undistorted even if a small number of bits
are changed in transit or on the storage media.
This property is used in, e.g., compact discs
(CDs).  The data on a CD can be reconstructed even
if the disc has been slightly scratched.

In a code-based encryption\index{encryption!code-based} scheme a message is
encrypted by adding a fixed number of errors to
(i.e. flipping a fixed numbers of bits of) the
encoded message.  Decoding requires knowledge of a
suitable decoding procedure which eliminates these
errors efficiently.  This method is the secret
key.  Code-based encryption is in general very
efficient.  At the moment, research focus on the
question which codes lead to secure encryption
schemes with keys which are as small as possible.

Encryption on the basis of lattice problems\index{encryption!lattice problems}
is very similar to that on the basis of
error-correcting codes.  Lattices are regular
structures of points in space.  For instance, the
points where the lines on squared paper cross form
a two-dimensional lattice.  For cryptographic
usage, the dimension of the lattices is chosen to
be much larger.  Encryption works as follows: The
plain-text is used to construct a lattice point
which is then slightly distorted in such a way
that it is no longer a lattice point, but close to
one.  Whoever knows a secret about the lattice is
able to find this lattice point in the vicinity of
the given point in space.  The lattice point in
turn yields the plain text.  A particularly
efficient lattice based encryption scheme is NTRU
Encrypt (\url{www.ntru.com})\index{encryption!lattice problems!NTR}.
However, because
NTRU was introduced fairly recently (in 1998), and
its specification underwent several changes due to
a variety of attacks, more cryptanalytic scrutiny
is required to achieve confidence in its security.

@@

In 1979, Ralph Merkle proposed a remarkable
framework for new signature schemes in his PhD
thesis~\cite{merkle-thesis:1979}.
Contrary to all other signature schemes\index{signature!Merkle}, its
security does not rest on the difficulty of a
number-theoretic, algebraic or geometric problem.
The only thing it requires is something which
other signature schemes need anyway: a
cryptographically secure hash function and a
secure pseudo-random number generator.  Each new
hash function leads to a new signature algorithm.
In consequence, the Merkle scheme has the
potential to solve the problem of long-term
availability of digital signature schemes.

@@

%\subsection{Quantum cryptography -- a loophole?}
\section{Quantum cryptography -- a way out of the impasse?}
\label{sec:quantum cryptography}\index{quantum cryptography}

@@

In order to guarantee IT security for the future, too,
we need to prepare a portfolio of secure cryptographic
schemes.  This portfolio needs to contain schemes
which are suitable for the world of ubiquitous
computing with many less powerful computers.  It
also needs to contain schemes which remain secure
in the event that powerful quantum computers\index{quantum computer} are
built.  Several promising candidates have been
discussed in this article.  They need to be
studied carefully and prepared for use in everyday
scenarios.  The question how to ensure long-term
confidentiality remains an important open research
problem upon which cryptographic research should
focus.
